Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-09-03 draft.tex
Filled in blanks in Document heading

\author{\authorname{Paul Klemm\sup{1}, Sylvia Glaßer\sup{1}, Kai Lawonn\sup{1} Henry Völzke\sup{2} and Bernhard Preim\sup{1}}
\affiliation{\sup{1}Department of Simulation and Graphics, University of Magdeburg}
\affiliation{\sup{2}Greifswald}
% TODO: Mailaddy nachtragen
\email{\{paul, sylvia\}@isg.cs.uni-magdeburg.de, voelzke@anderAddy}
}

% TODO: The paper must have at least one keyword. The text must be set to 9-point font size and without the use of bold or italic font style. For more than one keyword, please use a comma as a separator. Keywords must be titlecased.
\keywords{My Keywords in Title Case}

% \abstract{
% The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words. The text must be set to 9-point font size. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.}
\abstract{ 
The goal of epidemiology is to provide insight into disease causations.
%%
Hence, subject groups (cohorts) are analyzed to correlate the subjects' varying lifestyles and medical properties with diseases.
%%
Recently, these cohort studies comprise medical image data as well.
%%
We assess the predictive quality of image-derived variables for lower back pain.
%%
The evaluation is embedded in an interactive visual analysis (IVA) framework that is based on 2,540 segmented lumbar spine data sets.
%%
The segmentation results are evaluated and quantified by employing shape-describing variables, such as the spine canal course, its curvature and torsion.
%%
We analyze the predictive power of shape-describing variables for non-image variables, e.g., pain indicators.
%%
Therefore, we automatically extract a decision tree classifier for each non-image variable.
%%
Furthermore, we provide an IVA technique to compare the classifiers with interactive scatterplot visualizations.
%%
As a first result, we conclude that image-based variables are only sufficient for classification of lifestyle factors.
%%
A correlation between lumbar spine shape and lower back pain disease was not existent in the automatically trained classifiers.
%%
However, the presented approach is a valuable extension for the IVA of epidemiological data.
%%
Hence, correlations between non-image parameters were successfully detected and characterized.
}

>>>>

\author{\authorname{Paul Klemm\sup{1}, Sylvia Glaßer\sup{1}, Kai Lawonn\sup{1} Henry Völzke\sup{2}, Katrin Hegenscheid\sup{3} and Bernhard Preim\sup{1}}
\affiliation{\sup{1}Department of Simulation and Graphics, University of Magdeburg}
\affiliation{\sup{2} Institut for Community Medicine, University of Greifswald}
\affiliation{\sup{3} Institute for Diagnostic Radiology and Neuro-Radiology, University of Greifswald}
\email{\{paul, sylvia\}@isg.cs.uni-magdeburg.de, \{katrin.hegenscheid, voelzke\}@uni-greifswald.de}
}

\keywords{Epidemiology, Interactive Visual Analysis}

% \abstract{The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words.}
\abstract{ 
The goal of epidemiology is to provide insight into disease causations.
%%
Hence, subject groups (cohorts) are analyzed to correlate the subjects' varying lifestyles and medical properties with diseases.
%%
Recently, these cohort studies comprise medical image data as well.
%%
We assess the predictive quality of image-derived variables for lower back pain.
%%
The evaluation is embedded in an interactive visual analysis (IVA) framework that is based on 2,540 segmented lumbar spine data sets.
%%
The segmentation results are evaluated and quantified by employing shape-describing variables, such as the spine canal course, its curvature and torsion.
%%
We analyze the predictive power of shape-describing variables for non-image variables, e.g., pain indicators.
%%
Therefore, we automatically extract a decision tree classifier for each non-image variable.
%%
Furthermore, we provide an IVA technique to compare the classifiers with interactive scatterplot visualizations.
%%
As a first result, we conclude that image-based variables are only sufficient for classification of lifestyle factors.
%%
A correlation between lumbar spine shape and lower back pain disease was not existent in the automatically trained classifiers.
%%
However, the presented approach is a valuable extension for the IVA of epidemiological data.
%%
Hence, correlations between non-image parameters were successfully detected and characterized.
}
